# DS 202 Final Project: Movies Analysis

### Created by Ishan Patel, Ethan Mangelsdorf, Vinayak Trigunayat

## Information about our data set:

## Introduction:

## Data Cleaning:

## Genres:

## Directors:

## Runtime:

## Gross vs IMDb:

## Stars:

## The "Most Common" Movie:

## Potential Limitations:


#Final Project Analysis of Top IMDB Movies:



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


*Information about our Data Set:*

Our dataset is provided via courtesy of Harshit Shankhdhar, from Kaggle. It is titled 'IMDB Movies Dataset', and has 54.4k downloads, from which we can conclude that it is a reasonably popular dataset. The dataset can be accessed from the following link: https://www.kaggle.com/datasets/harshitshankhdhar/imdb-dataset-of-top-1000-movies-and-tv-shows. The dataset includes the top 1000 movies on IMDb as identified by their IMDb score. There are 16 columns and 1000 rows. Each row represents a separate film, and each column provides information about the film.


*Introduction:*

Through our analysis of the dataset, we aim to understand what variables are involved in the popularity of a movie, what are some ways to define popularity, and what some of the average values of the variables might be. 

*Cleaning: *

We start by importing the tidyverse package. We will also read in the dataset from the Kaggle csv file. We start by taking a look at the first few values.

```{r}
library(tidyverse)
imdb <- read.csv("imdb_top_1000.csv")
head(imdb)
```

We note that our analysis will not make use of the Poster_Link or Overview columns, so we can remove these. We then check for NA values. Since there are so few rows with NA values, we can comfortably remove these rows from our analysis.

```{r}
#Data Cleaning:

#Remove some columns
imdb <- imdb %>% select(-Poster_Link,-Overview)
results <- data.frame(name=rep(0,14),count=rep(0,14))
#Check how many NAs
for (column in 1:14) {
  #print(imdb[column])
  results$name[column] <- colnames(imdb[column])
  results$count[column] <- sum(is.na(imdb[column]))
}

#We observe that there are only NA values in the Meta_score.

```

The next part of analysis focuses on the Genre column.

*Topic 1: Genres* 

What are the highest rated genres?

```{r}
library(dplyr)
library(ggplot2)


# Split genres and count the frequency of each genre
genre_counts <- imdb %>%
  mutate(Genre = strsplit(Genre, ", ")) %>%
  unnest(Genre) %>%
  group_by(Genre) %>%
  summarize(Frequency = n()) %>%
  arrange(desc(Frequency))

# Reorder the Genre factor by Frequency
genre_counts$Genre <- factor(genre_counts$Genre, levels = genre_counts$Genre[order(genre_counts$Frequency, decreasing = TRUE)])

# Create a bar plot
ggplot(genre_counts, aes(x = Genre, y = Frequency, fill = Genre)) +
  geom_bar(stat = "identity") +
  labs(title = "Frequency of Genres in IMDb Dataset",
       x = "Genre",
       y = "Frequency") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))


```


This code first splits the genres listed for each movie, counts the frequency of each genre, and then creates a bar plot showing the frequency of each genre in the IMDb dataset. Drama is the most frequent.


What is the average IMDB Rating per Genre



```{r}
# Calculate the average IMDb rating for each genre
genre_avg_ratings <- imdb %>%
  mutate(Genre = strsplit(Genre, ", ")) %>%
  unnest(Genre) %>%
  group_by(Genre) %>%
  summarize(Avg_IMDB_Rating = mean(IMDB_Rating, na.rm = TRUE))

# Create a bar plot for average IMDb rating
ggplot(genre_avg_ratings, aes(x = Genre, y = Avg_IMDB_Rating, fill = Genre)) +
  geom_bar(stat = "identity") +
  labs(title = "Average IMDb Rating for Genres in IMDb Dataset",
       x = "Genre",
       y = "Average IMDb Rating") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```


This code calculates the average IMDb rating for each genre in the IMDb dataset. It first splits the genres listed for each movie, calculates the mean IMDb rating for each genre, and then creates a bar plot showing the average IMDb rating for each genre.

Western being the highest average IMDB score.


```{r}

library(knitr)

table <- kable(genre_avg_ratings, caption = "Average IMDb Rating for Genres in IMDb Dataset")

# Print the table
print(table)
```


*Topic 2: Directors*

Graph 1: Bar graph with all directors with their frequency 
```{r}


# Split directors and count the frequency of each director
director_counts <- imdb %>%
  group_by(Director) %>%
  summarize(Frequency = n()) %>%
  arrange(desc(Frequency))

# Select the top 10 directors
top_directors <- head(director_counts$Director, 10)

# Filter the dataset to include only the top 10 directors
top_director_counts <- director_counts %>%
  filter(Director %in% top_directors)

# Create a bar plot for the top 10 directors
ggplot(top_director_counts, aes(x = reorder(Director, Frequency), y = Frequency, fill = Director)) +
  geom_bar(stat = "identity") +
  labs(title = "Top 10 Most Frequent Directors in IMDb Dataset",
       x = "Director",
       y = "Frequency") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```



This code segment analyzes the frequency of directors in the IMDb dataset. It first counts the number of movies each director has made, then selects the top 10 most frequent directors. Finally, it creates a bar plot showing the frequency of movies directed by each of the top 10 directors

Alfred Hitchcock being the most frequent.


What are the top 3 directors in 1990’s?

```{r}


# Filter the dataset to include only movies from the 1990s
movies_1990s <- imdb %>%
  filter(Released_Year >= 1990 & Released_Year < 2000)

# Calculate the average IMDb score for each director
director_avg_ratings_1990s <- movies_1990s %>%
  mutate(Director = strsplit(Director, ", ")) %>%
  unnest(Director) %>%
  group_by(Director) %>%
  summarize(Avg_IMDB_Rating = mean(IMDB_Rating, na.rm = TRUE)) %>%
  arrange(desc(Avg_IMDB_Rating))

# Select the top 3 directors
top3_directors_1990s <- head(director_avg_ratings_1990s, 3)

# Create a bar plot for the top 3 directors
ggplot(top3_directors_1990s, aes(x = reorder(Director, Avg_IMDB_Rating), y = Avg_IMDB_Rating, fill = Director)) +
  geom_bar(stat = "identity") +
  labs(title = "Top 3 Directors in the 1990s by Average IMDb Rating",
       x = "Director",
       y = "Average IMDb Rating") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))



```


This code filters the IMDb dataset to include only movies released in the 1990s. It then calculates the average IMDb rating for each director based on their movies in the 1990s, selects the top 3 directors with the highest average ratings, and creates a bar plot showing their average ratings.




What are the top 3 directors in 2000’s?


```{r}


# Filter the dataset to include only movies from the 2000s
movies_2000s <- imdb %>%
  filter(Released_Year >= 2000 & Released_Year < 2010)

# Calculate the average IMDb score for each director
director_avg_ratings_2000s <- movies_2000s %>%
  mutate(Director = strsplit(Director, ", ")) %>%
  unnest(Director) %>%
  group_by(Director) %>%
  summarize(Avg_IMDB_Rating = mean(IMDB_Rating, na.rm = TRUE)) %>%
  arrange(desc(Avg_IMDB_Rating))

# Select the top 3 directors
top3_directors_2000s <- head(director_avg_ratings_2000s, 3)

# Create a bar plot for the top 3 directors
ggplot(top3_directors_2000s, aes(x = reorder(Director, Avg_IMDB_Rating), y = Avg_IMDB_Rating, fill = Director)) +
  geom_bar(stat = "identity") +
  labs(title = "Top 3 Directors in the 2000s by Average IMDb Rating",
       x = "Director",
       y = "Average IMDb Rating") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))



```

Same as before but for the 2000's



What are the top 3 directors in 2010’s?


```{r}
# Filter the dataset to include only movies from the 2010s
movies_2010s <- imdb %>%
  filter(Released_Year >= 2010 & Released_Year < 2020)

# Calculate the average IMDb score for each director
director_avg_ratings_2010s <- movies_2010s %>%
  mutate(Director = strsplit(Director, ", ")) %>%
  unnest(Director) %>%
  group_by(Director) %>%
  summarize(Avg_IMDB_Rating = mean(IMDB_Rating, na.rm = TRUE)) %>%
  arrange(desc(Avg_IMDB_Rating))

# Select the top 3 directors
top3_directors_2010s <- head(director_avg_ratings_2010s, 3)

# Create a bar plot for the top 3 directors
ggplot(top3_directors_2010s, aes(x = reorder(Director, Avg_IMDB_Rating), y = Avg_IMDB_Rating, fill = Director)) +
  geom_bar(stat = "identity") +
  labs(title = "Top 3 Directors in the 2010s by Average IMDb Rating",
       x = "Director",
       y = "Average IMDb Rating") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))



```

Same as before but for the 2010's

The number 1 highest scored director from each decade is :

Frank Darabont 1990's

Peter Jackson 2000's

Bong Joon Ho 2010's



```{r}
# Filter the dataset for each director
frank_darabont_movies <- imdb %>%
  filter(str_detect(Director, "Frank Darabont"))

peter_jackson_movies <- imdb %>%
  filter(str_detect(Director, "Peter Jackson"))

bong_joon_ho_movies <- imdb %>%
  filter(str_detect(Director, "Bong Joon Ho"))

# Count the number of movies for each director
frank_darabont_count <- nrow(frank_darabont_movies)
peter_jackson_count <- nrow(peter_jackson_movies)
bong_joon_ho_count <- nrow(bong_joon_ho_movies)

# Print the counts with custom messages
cat("Frank Darabont's amount of movies is", frank_darabont_count, "\n")
cat("Peter Jackson's amount of movies is", peter_jackson_count, "\n")
cat("Bong Joon Ho's amount of movies is", bong_joon_ho_count, "\n")

```

Frank Darabont's amount of movies is 2 
Peter Jackson's amount of movies is 5 
Bong Joon Ho's amount of movies is 3

Write each directors most common genre?

```{r}
# Function to find the most common genre for a director
find_most_common_genre <- function(director_movies) {
  genre_counts <- director_movies %>%
    mutate(Genre = strsplit(Genre, ", ")) %>%
    unnest(Genre) %>%
    group_by(Genre) %>%
    summarize(Frequency = n()) %>%
    arrange(desc(Frequency))
  
  most_common_genre <- genre_counts$Genre[1]
  return(most_common_genre)
}

# Find the most common genre for each director
frank_darabont_genre <- find_most_common_genre(frank_darabont_movies)
peter_jackson_genre <- find_most_common_genre(peter_jackson_movies)
bong_joon_ho_genre <- find_most_common_genre(bong_joon_ho_movies)

# Print the most common genres
cat("Frank Darabont's most common genre is", frank_darabont_genre, "\n")
cat("Peter Jackson's most common genre is", peter_jackson_genre, "\n")
cat("Bong Joon Ho's most common genre is", bong_joon_ho_genre, "\n")




```

Frank Darabont's most common genre is Drama 
Peter Jackson's most common genre is Adventure 
Bong Joon Ho's most common genre is Drama 

As you see two of each decade's best director most commonly made Drama Movies.



----------------


This scatter plot shows the correlation between gross revenue and IMDb ratings of all the movies in the dataset. This allows us to see if the gross revenue of a movie has an effect on its IMDb rating. From this scatter plot, we can see that there is a weak but positive correlation between the two variables.


```{r}
library(ggplot2)
library(dplyr)

# Convert Gross to numeric format and remove NA values
imdbGross <- imdb %>%
  mutate(Gross = as.numeric(gsub(",", "", Gross))) %>%
  filter(!is.na(Gross))

ggplot(imdbGross, aes(x = Gross, y = IMDB_Rating)) +
  geom_point(color = "skyblue", alpha = 0.6) +  # Scatter plot
  geom_smooth(method = "lm", se = FALSE, color = "red") +  # Add best-fit line
  labs(title = "Correlation Between Gross Earnings and IMDb Rating",
       x = "Gross Earnings", y = "IMDb Rating") +
  scale_x_continuous(labels = scales::comma) +
  theme_minimal()
```

This graph shows the frequency distribution of gross earnings from the dataset. We had to convert the “Gross” column to the numeric format as there originally were commas in the values. This graph is useful to see how frequently different ranges of gross earnings occur among the movies in the dataset. It also helps identify outliers and provide insights into successful and unsuccessful movies.


```{r}
library(ggplot2)

# Convert Gross to numeric format
imdb$Gross <- as.numeric(gsub(",", "", imdb$Gross))

# Define the bin width for the histogram
bin_width <- 50000000  # Adjust the bin width to suit the range of gross earnings

# Create the histogram
ggplot(imdb, aes(x = Gross)) +
  geom_histogram(binwidth = bin_width, fill = "skyblue", color = "black") +
  scale_x_continuous(labels = scales::comma) +  # Format x-axis labels with commas
  labs(title = "Frequency of Gross Earnings",
       x = "Gross Earnings", y = "Frequency") +
  theme_minimal()
```

This scatter plot helps us visualize the correlation between movie runtime and IMDb ratings. This correlation is slightly weak but positive. Here, we are looking at if a movie’s runtime influences its IMDb rating.

```{r}
library(ggplot2)
library(dplyr)

# Convert Runtime to numeric format and remove NA values
imdbRuntime <- imdb %>%
  mutate(Runtime = as.numeric(gsub(" min", "", Runtime))) %>%
  filter(!is.na(Runtime))

# Create the scatter plot
ggplot(imdbRuntime, aes(x = Runtime, y = IMDB_Rating)) +
  geom_point(color = "skyblue", alpha = 0.6) +  # Scatter plot
  geom_smooth(method = "lm", se = FALSE, color = "red") +  # Add best-fit line
  labs(title = "Correlation Between Runtime and IMDb Rating",
       x = "Runtime (minutes)", y = "IMDb Rating") +
  theme_minimal()
```

This graph provides insight into the distribution of movie runtimes in the dataset, allowing us to see how common different runtime lengths are among the movies. This will help us create different ranges of runtimes and see their average IMDb ratings, then we can see what would be an ideal runtime for a movie.

```{r}
# Convert Runtime to numeric format
imdb$Runtime <- as.numeric(gsub(" min", "", imdb$Runtime))

# Create the histogram for Runtime
ggplot(imdb, aes(x = Runtime)) +
  geom_histogram(binwidth = 10, fill = "skyblue", color = "black") +
  labs(title = "Histogram of Runtime",
       x = "Runtime (minutes)", y = "Frequency") +
  theme_minimal()
```
This graph visualizes the distribution of IMDb ratings and helps us identify the central tendency of IMDb ratings, indicating the most common or typical ratings. Peaks in the histogram suggest where the most ratings are concentrated, providing insight into the average or typical perception of movie quality among viewers.


```{r}
ggplot(imdb, aes(x = IMDB_Rating)) +
  geom_histogram(binwidth = 0.1, fill = "skyblue", color = "black") +
  labs(title = "Histogram of IMDb Rating",
       x = "IMDb Rating", y = "Frequency") +
  theme_minimal()
```
This box plot visualizes the distribution of IMDb ratings across different ranges of movie runtimes. This graph helps us identify the medians of each range which allows us to compare the medians across the different ranges. By comparing the box plots across different runtime ranges, we can assess whether there are significant differences in IMDb ratings based on movie runtimes.


```{r}
library(ggplot2)

# Convert Runtime to numeric format
imdb$Runtime <- as.numeric(gsub(" min", "", imdb$Runtime))

# Create the box plot
ggplot(imdb, aes(x = cut(Runtime, breaks = 10), y = IMDB_Rating)) +
  geom_boxplot(fill = "skyblue", color = "black") +
  labs(title = "Box Plot of IMDb Ratings by Runtime",
       x = "Runtime (minutes)", y = "IMDb Rating") +
  theme_minimal()
```

This graph identifies the top ten movies with the longest runtimes in the dataset. This information is valuable for identifying the top end of the spectrum.


```{r}
library(ggplot2)

# Assuming 'imdb' is your dataframe containing the entire dataset

# Convert 'Runtime' column to numeric format
imdb$Runtime <- as.numeric(gsub(" min", "", imdb$Runtime))

# Select the top ten movies with the longest runtime
top_ten_runtime <- imdb[order(imdb$Runtime, decreasing = TRUE), ][1:10, ]

# Create a bar chart for the top ten runtimes
ggplot(top_ten_runtime, aes(x = reorder(Series_Title, Runtime), y = Runtime)) +
  geom_bar(stat = "identity", fill = "skyblue") +
  labs(title = "Top Ten Runtimes",
       x = "Movie",
       y = "Runtime (minutes)") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  coord_flip()
```

This table orders the top ten movies with the longest runtimes by their IMDb ratings in descending order. This will help us see what the top end of the spectrum regarding runtime is being rated. We will be able to see whether audiences prefer longer movies over shorter ones or vice versa.


```{r}

top_ten_runtime <- imdb[order(imdb$Runtime, decreasing = TRUE), ][1:10, ]

top_ten_ratings <- top_ten_runtime$IMDB_Rating



top_ten_table <- data.frame(Movie = top_ten_runtime$Series_Title, IMDb_Rating = top_ten_ratings)

# Order the table by IMDb ratings in descending order
top_ten_table <- top_ten_table[order(top_ten_table$IMDb_Rating, decreasing = TRUE), ]

# Print the table
print(top_ten_table)

average_rating <- mean(top_ten_ratings)

# Print the average rating
print(average_rating)
```

Like the top ten runtimes, we look at the other end of the spectrum regarding runtime length.

```{r}
bottom_ten_runtime <- imdb[order(imdb$Runtime, decreasing = FALSE), ][1:10, ]

# Create a bar chart for the bottom ten runtimes
ggplot(bottom_ten_runtime, aes(x = reorder(Series_Title, Runtime), y = Runtime)) +
  geom_bar(stat = "identity", fill = "red") +
  labs(title = "Bottom Ten Runtimes",
       x = "Movie",
       y = "Runtime (minutes)") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  coord_flip()
```

This table orders the bottom ten movies with the longest runtimes by their IMDb ratings in descending order. This will help us see what the bottom end of the spectrum regarding runtime is being rated. We can also see the average of the bottom ten movies and compare that to the average of the top ten.

```{r}
bottom_ten_runtime <- imdb[order(imdb$Runtime, decreasing = FALSE), ][1:10, ]

# Subset IMDb ratings for the bottom ten movies
bottom_ten_ratings <- bottom_ten_runtime$IMDB_Rating

# Create a dataframe with movie titles and IMDb ratings
bottom_ten_table <- data.frame(Movie = bottom_ten_runtime$Series_Title, IMDb_Rating = bottom_ten_ratings)

# Order the table by IMDb ratings in descending order
bottom_ten_table <- bottom_ten_table[order(bottom_ten_table$IMDb_Rating, decreasing = TRUE), ]

# Print the table
print(bottom_ten_table)
average_rating <- mean(bottom_ten_ratings)

# Print the average rating
print(average_rating)
```

Like the top and bottom ten runtimes, we look at the top ten grossing movies to dive deeper into the correlation between the gross revenue and IMDb rating.


```{r}
library(scales)

# Create a bar chart for the top ten grossing movies with formatted y-axis labels
ggplot(top_ten_grossing, aes(x = reorder(Series_Title, Gross), y = Gross)) +
  geom_bar(stat = "identity", fill = "forestgreen") +
  labs(title = "Top Ten Grossing Movies",
       x = "Movie",
       y = "Gross Earnings") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_y_continuous(labels = scales::comma, breaks = c(0, 250000000, 500000000, 750000000)) +
  coord_flip()
```

This table organizes the top ten grossing movies in descending order based on their IMDb ratings and we can get an average IMDb rating for the 10 movies and compare that to the overall average and the bottom ten average.

```{r}
# Select the top ten movies with the highest gross earnings
top_ten_grossing <- imdb[order(imdb$Gross, decreasing = TRUE), ][1:10, ]

# Create a dataframe with movie titles and IMDb ratings
top_ten_ratings_table <- data.frame(Movie = top_ten_grossing$Series_Title, IMDb_Rating = top_ten_grossing$IMDB_Rating)

# Order the table by IMDb ratings in descending order
top_ten_ratings_table <- top_ten_ratings_table[order(top_ten_ratings_table$IMDb_Rating, decreasing = TRUE), ]

# Print the table
print(top_ten_ratings_table)
```

This graph identifies the bottom ten grossing movies and their revenues.

```{r}
# Select the bottom ten movies with the lowest gross earnings
bottom_ten_grossing <- imdb[order(imdb$Gross), ][1:10, ]

# Create a bar chart for the bottom ten grossing movies
ggplot(bottom_ten_grossing, aes(x = reorder(Series_Title, Gross), y = Gross)) +
  geom_bar(stat = "identity", fill = "firebrick") +
  labs(title = "Bottom Ten Grossing Movies",
       x = "Movie",
       y = "Gross Earnings") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  coord_flip()
```

This table orders the bottom ten grossing movies by their IMDb ratings in descending order. The table will allow us to compare the average IMDb ratings of the two ends of the gross revenue spectrum to see if a higher-grossing movie will, on average, have a higher IMDb rating than a lower-grossing movie.


```{r}
bottom_ten_grossing <- imdb[order(imdb$Gross), ][1:10, ]

# Create a dataframe with movie titles and IMDb ratings
bottom_ten_ratings_table <- data.frame(Movie = bottom_ten_grossing$Series_Title, IMDb_Rating = bottom_ten_grossing$IMDB_Rating)

# Order the table by IMDb ratings in descending order
bottom_ten_ratings_table <- bottom_ten_ratings_table[order(bottom_ten_ratings_table$IMDb_Rating, decreasing = TRUE), ]

# Print the table
print(bottom_ten_ratings_table)

```


*Stars:*









