# DS 202 Final Project: Movies Analysis

### Created by Ishan Patel, Ethan Mangelsdorf, Vinayak Trigunayat

## Information about our data set:

## Introduction:

## Data Cleaning:

## Genres:

## Directors:

## Runtime:

## Gross vs IMDb:

## Stars:

## The "Most Common" Movie:

## Potential Limitations:


#Final Project Analysis of Top IMDB Movies:


<<<<<<< HEAD

=======
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

*Introduction:*

This project intends to analyse the features of the top IMDb movies of all time. 



```{r}
library(tidyverse)
imdb <- read.csv("imdb_top_1000.csv")
head(imdb)
```

*Cleaning: *
```{r}
#Data Cleaning:

#Remove some columns
imdb <- imdb %>% select(-Poster_Link,-Overview)
results <- data.frame(name=rep(0,14),count=rep(0,14))
#Check how many NAs
for (column in 1:14) {
  #print(imdb[column])
  results$name[column] <- colnames(imdb[column])
  results$count[column] <- sum(is.na(imdb[column]))
}

#We observe that there are only NA values in the Meta_score.

 # imdb2 <- imdb %>% separate(`Genre`, sep=",", into)

```
>>>>>>> 83ba0457a9196ba3b477d38a949e5006d4513c72



Topic 1: Genres 

What are the highest rated genres?

```{r}
library(dplyr)
library(ggplot2)


# Split genres and count the frequency of each genre
genre_counts <- imdb %>%
  mutate(Genre = strsplit(Genre, ", ")) %>%
  unnest(Genre) %>%
  group_by(Genre) %>%
  summarize(Frequency = n()) %>%
  arrange(desc(Frequency))

# Reorder the Genre factor by Frequency
genre_counts$Genre <- factor(genre_counts$Genre, levels = genre_counts$Genre[order(genre_counts$Frequency, decreasing = TRUE)])

# Create a bar plot
ggplot(genre_counts, aes(x = Genre, y = Frequency, fill = Genre)) +
  geom_bar(stat = "identity") +
  labs(title = "Frequency of Genres in IMDb Dataset",
       x = "Genre",
       y = "Frequency") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))


```


What is the average IMDB Rating per Genre



```{r}
# Calculate the average IMDb rating for each genre
genre_avg_ratings <- imdb %>%
  mutate(Genre = strsplit(Genre, ", ")) %>%
  unnest(Genre) %>%
  group_by(Genre) %>%
  summarize(Avg_IMDB_Rating = mean(IMDB_Rating, na.rm = TRUE))

# Create a bar plot for average IMDb rating
ggplot(genre_avg_ratings, aes(x = Genre, y = Avg_IMDB_Rating, fill = Genre)) +
  geom_bar(stat = "identity") +
  labs(title = "Average IMDb Rating for Genres in IMDb Dataset",
       x = "Genre",
       y = "Average IMDb Rating") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```




```{r}

library(knitr)

table <- kable(genre_avg_ratings, caption = "Average IMDb Rating for Genres in IMDb Dataset")

# Print the table
print(table)
```






Topic 2 Director Ethan 

Graph 1: Bar graph with all directors with their frequency 
```{r}

# Assuming 'imdb' is your dataset
# Split directors and count the frequency of each director
director_counts <- imdb %>%
  group_by(Director) %>%
  summarize(Frequency = n()) %>%
  arrange(desc(Frequency))

# Select the top 10 directors
top_directors <- head(director_counts$Director, 10)

# Filter the dataset to include only the top 10 directors
top_director_counts <- director_counts %>%
  filter(Director %in% top_directors)

# Create a bar plot for the top 10 directors
ggplot(top_director_counts, aes(x = reorder(Director, Frequency), y = Frequency, fill = Director)) +
  geom_bar(stat = "identity") +
  labs(title = "Top 10 Most Frequent Directors in IMDb Dataset",
       x = "Director",
       y = "Frequency") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```





What are the top 3 directors in 1990’s?

```{r}


# Filter the dataset to include only movies from the 1990s
movies_1990s <- imdb %>%
  filter(Released_Year >= 1990 & Released_Year < 2000)

# Calculate the average IMDb score for each director
director_avg_ratings_1990s <- movies_1990s %>%
  mutate(Director = strsplit(Director, ", ")) %>%
  unnest(Director) %>%
  group_by(Director) %>%
  summarize(Avg_IMDB_Rating = mean(IMDB_Rating, na.rm = TRUE)) %>%
  arrange(desc(Avg_IMDB_Rating))

# Select the top 3 directors
top3_directors_1990s <- head(director_avg_ratings_1990s, 3)

# Create a bar plot for the top 3 directors
ggplot(top3_directors_1990s, aes(x = reorder(Director, Avg_IMDB_Rating), y = Avg_IMDB_Rating, fill = Director)) +
  geom_bar(stat = "identity") +
  labs(title = "Top 3 Directors in the 1990s by Average IMDb Rating",
       x = "Director",
       y = "Average IMDb Rating") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))



```







What are the top 3 directors in 2000’s?


```{r}


# Filter the dataset to include only movies from the 2000s
movies_2000s <- imdb %>%
  filter(Released_Year >= 2000 & Released_Year < 2010)

# Calculate the average IMDb score for each director
director_avg_ratings_2000s <- movies_2000s %>%
  mutate(Director = strsplit(Director, ", ")) %>%
  unnest(Director) %>%
  group_by(Director) %>%
  summarize(Avg_IMDB_Rating = mean(IMDB_Rating, na.rm = TRUE)) %>%
  arrange(desc(Avg_IMDB_Rating))

# Select the top 3 directors
top3_directors_2000s <- head(director_avg_ratings_2000s, 3)

# Create a bar plot for the top 3 directors
ggplot(top3_directors_2000s, aes(x = reorder(Director, Avg_IMDB_Rating), y = Avg_IMDB_Rating, fill = Director)) +
  geom_bar(stat = "identity") +
  labs(title = "Top 3 Directors in the 2000s by Average IMDb Rating",
       x = "Director",
       y = "Average IMDb Rating") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))



```





What are the top 3 directors in 2010’s?


```{r}
# Filter the dataset to include only movies from the 2010s
movies_2010s <- imdb %>%
  filter(Released_Year >= 2010 & Released_Year < 2020)

# Calculate the average IMDb score for each director
director_avg_ratings_2010s <- movies_2010s %>%
  mutate(Director = strsplit(Director, ", ")) %>%
  unnest(Director) %>%
  group_by(Director) %>%
  summarize(Avg_IMDB_Rating = mean(IMDB_Rating, na.rm = TRUE)) %>%
  arrange(desc(Avg_IMDB_Rating))

# Select the top 3 directors
top3_directors_2010s <- head(director_avg_ratings_2010s, 3)

# Create a bar plot for the top 3 directors
ggplot(top3_directors_2010s, aes(x = reorder(Director, Avg_IMDB_Rating), y = Avg_IMDB_Rating, fill = Director)) +
  geom_bar(stat = "identity") +
  labs(title = "Top 3 Directors in the 2010s by Average IMDb Rating",
       x = "Director",
       y = "Average IMDb Rating") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))



```


The number 1 highest scored director from each decade is :

Frank Darabont 1990's

Peter Jackson 2000's

Bong Joon Ho 2010's



```{r}
# Filter the dataset for each director
frank_darabont_movies <- imdb %>%
  filter(str_detect(Director, "Frank Darabont"))

peter_jackson_movies <- imdb %>%
  filter(str_detect(Director, "Peter Jackson"))

bong_joon_ho_movies <- imdb %>%
  filter(str_detect(Director, "Bong Joon Ho"))

# Count the number of movies for each director
frank_darabont_count <- nrow(frank_darabont_movies)
peter_jackson_count <- nrow(peter_jackson_movies)
bong_joon_ho_count <- nrow(bong_joon_ho_movies)

# Print the counts with custom messages
cat("Frank Darabont's amount of movies is", frank_darabont_count, "\n")
cat("Peter Jackson's amount of movies is", peter_jackson_count, "\n")
cat("Bong Joon Ho's amount of movies is", bong_joon_ho_count, "\n")

```



Write each directors most common genre?

```{r}
# Function to find the most common genre for a director
find_most_common_genre <- function(director_movies) {
  genre_counts <- director_movies %>%
    mutate(Genre = strsplit(Genre, ", ")) %>%
    unnest(Genre) %>%
    group_by(Genre) %>%
    summarize(Frequency = n()) %>%
    arrange(desc(Frequency))
  
  most_common_genre <- genre_counts$Genre[1]
  return(most_common_genre)
}

# Find the most common genre for each director
frank_darabont_genre <- find_most_common_genre(frank_darabont_movies)
peter_jackson_genre <- find_most_common_genre(peter_jackson_movies)
bong_joon_ho_genre <- find_most_common_genre(bong_joon_ho_movies)

# Print the most common genres
cat("Frank Darabont's most common genre is", frank_darabont_genre, "\n")
cat("Peter Jackson's most common genre is", peter_jackson_genre, "\n")
cat("Bong Joon Ho's most common genre is", bong_joon_ho_genre, "\n")




```


Are there any correlation in the top directors genre?


As you see two of each decade's best director most commonly made Drama Movies.



----------------


```{r}
library(ggplot2)
library(dplyr)

# Convert Gross to numeric format and remove NA values
imdbGross <- imdb %>%
  mutate(Gross = as.numeric(gsub(",", "", Gross))) %>%
  filter(!is.na(Gross))

ggplot(imdbGross, aes(x = Gross, y = IMDB_Rating)) +
  geom_point(color = "skyblue", alpha = 0.6) +  # Scatter plot
  geom_smooth(method = "lm", se = FALSE, color = "red") +  # Add best-fit line
  labs(title = "Correlation Between Gross Earnings and IMDb Rating",
       x = "Gross Earnings", y = "IMDb Rating") +
  scale_x_continuous(labels = scales::comma) +
  theme_minimal()
```

```{r}
library(ggplot2)

# Convert Gross to numeric format
imdb$Gross <- as.numeric(gsub(",", "", imdb$Gross))

# Define the bin width for the histogram
bin_width <- 50000000  # Adjust the bin width to suit the range of gross earnings

# Create the histogram
ggplot(imdb, aes(x = Gross)) +
  geom_histogram(binwidth = bin_width, fill = "skyblue", color = "black") +
  scale_x_continuous(labels = scales::comma) +  # Format x-axis labels with commas
  labs(title = "Frequency of Gross Earnings",
       x = "Gross Earnings", y = "Frequency") +
  theme_minimal()
```

```{r}
library(ggplot2)
library(dplyr)

# Convert Runtime to numeric format and remove NA values
imdbRuntime <- imdb %>%
  mutate(Runtime = as.numeric(gsub(" min", "", Runtime))) %>%
  filter(!is.na(Runtime))

# Create the scatter plot
ggplot(imdbRuntime, aes(x = Runtime, y = IMDB_Rating)) +
  geom_point(color = "skyblue", alpha = 0.6) +  # Scatter plot
  geom_smooth(method = "lm", se = FALSE, color = "red") +  # Add best-fit line
  labs(title = "Correlation Between Runtime and IMDb Rating",
       x = "Runtime (minutes)", y = "IMDb Rating") +
  theme_minimal()
```

```{r}
# Convert Runtime to numeric format
imdb$Runtime <- as.numeric(gsub(" min", "", imdb$Runtime))

# Create the histogram for Runtime
ggplot(imdb, aes(x = Runtime)) +
  geom_histogram(binwidth = 10, fill = "skyblue", color = "black") +
  labs(title = "Histogram of Runtime",
       x = "Runtime (minutes)", y = "Frequency") +
  theme_minimal()
```

```{r}
ggplot(imdb, aes(x = IMDB_Rating)) +
  geom_histogram(binwidth = 0.1, fill = "skyblue", color = "black") +
  labs(title = "Histogram of IMDb Rating",
       x = "IMDb Rating", y = "Frequency") +
  theme_minimal()
```

```{r}
library(ggplot2)

# Convert Runtime to numeric format
imdb$Runtime <- as.numeric(gsub(" min", "", imdb$Runtime))

# Create the box plot
ggplot(imdb, aes(x = cut(Runtime, breaks = 10), y = IMDB_Rating)) +
  geom_boxplot(fill = "skyblue", color = "black") +
  labs(title = "Box Plot of IMDb Ratings by Runtime",
       x = "Runtime (minutes)", y = "IMDb Rating") +
  theme_minimal()
```

```{r}
library(ggplot2)

# Assuming 'imdb' is your dataframe containing the entire dataset

# Convert 'Runtime' column to numeric format
imdb$Runtime <- as.numeric(gsub(" min", "", imdb$Runtime))

# Select the top ten movies with the longest runtime
top_ten_runtime <- imdb[order(imdb$Runtime, decreasing = TRUE), ][1:10, ]

# Create a bar chart for the top ten runtimes
ggplot(top_ten_runtime, aes(x = reorder(Series_Title, Runtime), y = Runtime)) +
  geom_bar(stat = "identity", fill = "skyblue") +
  labs(title = "Top Ten Runtimes",
       x = "Movie",
       y = "Runtime (minutes)") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  coord_flip()
```

```{r}

top_ten_runtime <- imdb[order(imdb$Runtime, decreasing = TRUE), ][1:10, ]

top_ten_ratings <- top_ten_runtime$IMDB_Rating



top_ten_table <- data.frame(Movie = top_ten_runtime$Series_Title, IMDb_Rating = top_ten_ratings)

# Order the table by IMDb ratings in descending order
top_ten_table <- top_ten_table[order(top_ten_table$IMDb_Rating, decreasing = TRUE), ]

# Print the table
print(top_ten_table)

average_rating <- mean(top_ten_ratings)

# Print the average rating
print(average_rating)
```

```{r}
bottom_ten_runtime <- imdb[order(imdb$Runtime, decreasing = FALSE), ][1:10, ]

# Create a bar chart for the bottom ten runtimes
ggplot(bottom_ten_runtime, aes(x = reorder(Series_Title, Runtime), y = Runtime)) +
  geom_bar(stat = "identity", fill = "red") +
  labs(title = "Bottom Ten Runtimes",
       x = "Movie",
       y = "Runtime (minutes)") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  coord_flip()
```

```{r}
bottom_ten_runtime <- imdb[order(imdb$Runtime, decreasing = FALSE), ][1:10, ]

# Subset IMDb ratings for the bottom ten movies
bottom_ten_ratings <- bottom_ten_runtime$IMDB_Rating

# Create a dataframe with movie titles and IMDb ratings
bottom_ten_table <- data.frame(Movie = bottom_ten_runtime$Series_Title, IMDb_Rating = bottom_ten_ratings)

# Order the table by IMDb ratings in descending order
bottom_ten_table <- bottom_ten_table[order(bottom_ten_table$IMDb_Rating, decreasing = TRUE), ]

# Print the table
print(bottom_ten_table)
average_rating <- mean(bottom_ten_ratings)

# Print the average rating
print(average_rating)
```

```{r}
library(scales)

# Create a bar chart for the top ten grossing movies with formatted y-axis labels
ggplot(top_ten_grossing, aes(x = reorder(Series_Title, Gross), y = Gross)) +
  geom_bar(stat = "identity", fill = "forestgreen") +
  labs(title = "Top Ten Grossing Movies",
       x = "Movie",
       y = "Gross Earnings") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_y_continuous(labels = scales::comma, breaks = c(0, 250000000, 500000000, 750000000)) +
  coord_flip()
```

```{r}
# Select the top ten movies with the highest gross earnings
top_ten_grossing <- imdb[order(imdb$Gross, decreasing = TRUE), ][1:10, ]

# Create a dataframe with movie titles and IMDb ratings
top_ten_ratings_table <- data.frame(Movie = top_ten_grossing$Series_Title, IMDb_Rating = top_ten_grossing$IMDB_Rating)

# Order the table by IMDb ratings in descending order
top_ten_ratings_table <- top_ten_ratings_table[order(top_ten_ratings_table$IMDb_Rating, decreasing = TRUE), ]

# Print the table
print(top_ten_ratings_table)
```

```{r}
# Select the bottom ten movies with the lowest gross earnings
bottom_ten_grossing <- imdb[order(imdb$Gross), ][1:10, ]

# Create a bar chart for the bottom ten grossing movies
ggplot(bottom_ten_grossing, aes(x = reorder(Series_Title, Gross), y = Gross)) +
  geom_bar(stat = "identity", fill = "firebrick") +
  labs(title = "Bottom Ten Grossing Movies",
       x = "Movie",
       y = "Gross Earnings") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  coord_flip()
```

```{r}
bottom_ten_grossing <- imdb[order(imdb$Gross), ][1:10, ]

# Create a dataframe with movie titles and IMDb ratings
bottom_ten_ratings_table <- data.frame(Movie = bottom_ten_grossing$Series_Title, IMDb_Rating = bottom_ten_grossing$IMDB_Rating)

# Order the table by IMDb ratings in descending order
bottom_ten_ratings_table <- bottom_ten_ratings_table[order(bottom_ten_ratings_table$IMDb_Rating, decreasing = TRUE), ]

# Print the table
print(bottom_ten_ratings_table)

```










